---
title: "ACCORD Analysis"
date: "2024-03-25"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)

expr1 <- read_csv("expr1/2024-10-9-16:40.csv")
expr2 <- read_csv("expr2/2024-10-21-11:50.csv")
expr3 <- read_csv("expr3/2024-11-08-10:05.csv")

expr1$std_dev <- apply(subset(expr1, select=c("dtime0","dtime1","dtime2","dtime3","dtime4","dtime5","dtime6","dtime7","dtime8","dtime9")), 1, sd)

expr2$log_file <- as.factor(expr2$log_file)
levels(expr2$log_file) <- c("Log 1", "Log 2", "Log 3", "Log 4")
log_files <- unique(expr2$log_file)
expr2.wider <- expr2 %>% select(!c("users","resources"))
expr2.wider <- expr2.wider %>% pivot_wider(id_cols=c(activity_count, selectivity_level), names_from=log_file, values_from=detection_time)
expr2.wider$avg_detection_time <- rowMeans(subset(expr2.wider, select = log_files))
expr2.wider$std_dev <- apply(subset(expr2.wider, select = log_files), 1, sd)


# write.csv(expr2.wider, "expr2-aggregated.csv")
```

## Experiment 1: Construction Time
```{r, echo=FALSE}
# expr1$std_dev
# summary(expr1)
ggplot(data=expr1, aes(x=constraint_count, y=dtime_avg)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=dtime_avg-std_dev, ymax=dtime_avg+std_dev)) +
  theme_bw() +
  labs(title="Construction time vs. action constraints", x="Action constraints", y="Construction time (ms)")
```

## Experiment 2
### Graph
```{r, echo=FALSE}
# summary(expr2.wider)
ggplot(data=expr2.wider, aes(x=activity_count, y=avg_detection_time, group=selectivity_level, color=selectivity_level)) +
  geom_line(aes(linetype=selectivity_level))+
  scale_linetype_manual(values=c("dashed", "twodash", "solid"), breaks=c('low', 'medium', 'high'), labels=c("Low >=20%", "Medium >=5% and <20%", "High <5%")) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=20%", "Medium >=5% and <20%", "High <5%")) +
  geom_point(show.legend=FALSE) +
  geom_errorbar(show.legend = FALSE, aes(ymin=avg_detection_time-std_dev, ymax=avg_detection_time+std_dev, width=25)) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Total time vs. activities by selectivity level", x="Activities", y="Detection time (ms)", color="Selectivity", linetype="Selectivity")

ggplot(data=expr2, aes(x=resources, y=detection_time, group=selectivity_level, color=selectivity_level)) +
  geom_point(aes(shape=log_file), breaks=c('low', 'medium', 'high'), labels=c("Low >=20%", "Medium >=5% and <20%", "High <5%")) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=20%", "Medium >=5% and <20%", "High <5%")) +
  theme_bw() +
  # theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Total time vs. resources by selectivity level", x="Resources", y="Detection time (ms)", color="Selectivity", shape="Log file")

ggplot(data=expr2, aes(x=5*resources*(3+2+3*5), y=detection_time, group=selectivity_level, color=selectivity_level)) +
  geom_point(breaks=c('low', 'medium', 'high'), labels=c("Low >=20%", "Medium >=5% and <20%", "High <5%")) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=20%", "Medium >=5% and <20%", "High <5%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Total time vs. action space size by selectivity level", x="Action space size", y="Detection time (ms)", color="Selectivity", shape="Log file")
```


### Stats
```{r, echo=FALSE}
ggplot(data=expr2, aes(x=selectivity_level, y=selectivity, fill=selectivity_level)) +
  geom_boxplot() +
  labs(title = "Variation of selectivity within a level", x = "Selectivity level", y = "Selectivity values")
```

Low selectivity
```{r, echo=FALSE}
expr2.low <- expr2 %>% filter(selectivity_level == "low")
expr2.low %>% summarise(mean = mean(selectivity),
            median = median(selectivity),
            std_dev = sd(selectivity),
            IQR = IQR(selectivity),
            min = min(selectivity),
            max = max(selectivity),
            var = var(selectivity),
            n = n())
```

Medium selectivity
```{r, echo=FALSE}
expr2.medium <- expr2 %>% filter(selectivity_level == "medium")
expr2.medium %>% summarise(mean = mean(selectivity),
            median = median(selectivity),
            std_dev = sd(selectivity),
            IQR = IQR(selectivity),
            min = min(selectivity),
            max = max(selectivity),
            var = var(selectivity),
            n = n())
```

High selectivity
```{r, echo=FALSE}
expr2.high <- expr2 %>% filter(selectivity_level == "high")
expr2.high %>% summarise(mean = mean(selectivity),
            median = median(selectivity),
            std_dev = sd(selectivity),
            IQR = IQR(selectivity),
            min = min(selectivity),
            max = max(selectivity),
            var = var(selectivity),
            n = n())

```

## Experiment 3
### Graphs for total run time (detection time + construction time)
```{r, echo=FALSE}
ggplot(data=expr3, aes(x=activity_count, y=detection_time_mean+construction_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean+construction_time_mean-(detection_time_std+construction_time_std), ymax=detection_time_mean+construction_time_mean+construction_time_std+detection_time_std, width=50)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Total run time vs. activities by selectivity level", x="Activities", y="Total run time (ms)", color="Selectivity")

ggplot(data=expr3, aes(x=users, y=detection_time_mean+construction_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean+construction_time_mean-(detection_time_std+construction_time_std), ymax=detection_time_mean+construction_time_mean+construction_time_std+detection_time_std, width=0.5)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Total run time vs. users by selectivity level", x="Users", y="Total run time (ms)", color="Selectivity")

ggplot(data=expr3, aes(x=resources, y=detection_time_mean+construction_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean+construction_time_mean-(detection_time_std+construction_time_std), ymax=detection_time_mean+construction_time_mean+construction_time_std+detection_time_std, width=10)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Total run time vs. resources by selectivity level", x="Resources", y="Total run time (ms)", color="Selectivity")
```

```{r, echo=FALSE}
ggplot(data=expr3, aes(x=resources*users*(5+3*users), y=construction_time_mean+detection_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=construction_time_mean+detection_time_mean-construction_time_std-detection_time_std, ymax=construction_time_mean+detection_time_mean+detection_time_std+construction_time_std, width=75000)) +
  theme_bw() +
  theme(legend.position="bottom") +
  labs(title="Total run time vs. action space size", x="Action space size", y="Total run time time", color="Selectivity level")
```

### Graphs for construction time
In the first graph, it appears that increasing the activity count will increase the construction time. However, since no logs are used in this portion of the experiment, this is impossible.
Rather, the activity count has a strong correlation with the action space size,
which is a strong predictor of the construction time.

```{r, echo=FALSE}
ggplot(data=expr3, aes(x=activity_count, y=construction_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=construction_time_mean-(construction_time_std), ymax=construction_time_mean+construction_time_std, width=50)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Construction time vs. activities by selectivity level", x="Activities", y="Construction time (ms)", color="Selectivity")

ggplot(data=expr3, aes(x=users, y=construction_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=construction_time_mean-(construction_time_std), ymax=construction_time_mean+construction_time_std, width=0.5)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Construction time vs. users by selectivity level", x="Users", y="Construction time (ms)", color="Selectivity")

ggplot(data=expr3, aes(x=resources, y=construction_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=construction_time_mean-(construction_time_std), ymax=construction_time_mean+construction_time_std, width=10)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Construction time vs. resources by selectivity level", x="Resources", y="Construction time (ms)", color="Selectivity")

ggplot(data=expr3, aes(x=resources*users*(5+3*users), y=construction_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=construction_time_mean-construction_time_std, ymax=construction_time_mean+construction_time_std, width=75000)) +
  theme_bw() +
  theme(legend.position="bottom") +
  labs(title="Construction time vs. action space size", x="Action space size", y="Construction time (ms)", color="Selectivity level")
```

### Graphs for detection time
```{r, echo=FALSE}
ggplot(data=expr3, aes(x=activity_count, y=detection_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean-(detection_time_std), ymax=detection_time_mean+detection_time_std, width=50)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Detection time vs. activities by selectivity level", x="Activities", y="Detection time (ms)", color="Selectivity")

ggplot(data=expr3, aes(x=users, y=detection_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean-(detection_time_std), ymax=detection_time_mean+detection_time_std, width=0.5)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Detection time vs. users by selectivity level", x="Users", y="Detection time (ms)", color="Selectivity")

ggplot(data=expr3, aes(x=resources, y=detection_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean-(detection_time_std), ymax=detection_time_mean+detection_time_std, width=10)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Detection time vs. resources by selectivity level", x="Actions", y="Detection time (ms)", color="Selectivity")


ggplot(data=expr3, aes(x=resources*users*(5+3*users), y=detection_time_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean-detection_time_std, ymax=detection_time_mean+detection_time_std, width=50000)) +
  theme_bw() +
  theme(legend.position="bottom") +
  labs(title="Detection time vs. action space size", x="Action space size", y="Detection time", color="Selectivity level")
```

### Construction time as a fraction of total time
```{r, echo=FALSE}
expr3.med <- expr3 %>% filter(selectivity_level == "medium")
ggplot(data=expr3.med, aes(x=resources*users*(5+3*users))) +
  geom_line(aes(y=construction_time_mean, color="Construction")) +
  geom_line(aes(y=construction_time_mean+detection_time_mean, color="Total")) +
  geom_point(aes(y=construction_time_mean, color="Construction")) +
  geom_point(aes(y=construction_time_mean+detection_time_mean, color="Total")) +
  theme(legend.title=element_blank()) +
  theme_bw() + 
  labs(y="Time (ms)", x="Action space size", title="Total run time and Construction time vs. action space size", colour="")
```

### Selectivity vs. Activity count
```{r, echo=FALSE}
ggplot(data=expr3, aes(x=activity_count, y=selectivity_mean, group=selectivity_level, color=selectivity_level)) +
  geom_point() +
  geom_errorbar(show.legend=FALSE,
                aes(ymin=selectivity_mean-selectivity_std, ymax=selectivity_mean+selectivity_std, width=50)) +
  scale_color_discrete(breaks=c('low', 'medium', 'high'), labels=c("Low >=30% and <40%", "Medium >=20% and <30%", "High >=10% and <20%")) +
  theme_bw() +
  theme(legend.position="bottom", legend.title=element_blank()) +
  labs(title="Selectivity vs. activities by selectivity level", x="Actions", y="Selectivity", color="Selectivity")
```

### Communicating how selectivity is also varying
Within a selectivity range, the lowest points have fewer activities, users, and resources.
If we keep this type of graph, I'd do the experiment again without using specific ranges
to eliminate the patterns in the data.

```{r, echo=FALSE}
ggplot(data=expr3, aes(x=selectivity_mean, y=detection_time_mean, group=selectivity_level)) +
  geom_point(aes(size=activity_count)) +
  geom_errorbar(show.legend = FALSE, aes(ymin=detection_time_mean-detection_time_std, ymax=detection_time_mean+detection_time_std, width=.01)) +
  theme_bw() +
  theme(legend.position="bottom") +
  labs(title="Detection time vs. selectivity", x="Selectivity", y="Detection time", size="Activity count")
```

```{r, echo=FALSE}
ggplot(data=expr3, aes(x=selectivity_mean, y=construction_time_mean, group=selectivity_level)) +
  geom_point(aes(size=resources*users*(5+3*users))) +
  geom_errorbar(show.legend = FALSE, aes(ymin=construction_time_mean-construction_time_std, ymax=construction_time_mean+construction_time_std, width=.01)) +
  theme_bw() +
  theme(legend.position="bottom") +
  labs(title="Construction time vs. selectivity", x="Selectivity", y="Construction time", size="Action space size")
```

### Regression
*Note: Assumptions that must be met before doing multiple linear regression weren't fully checked.
The construction time model for sure is inaccurate, because this has been demonstrated to not be a linear relationship. Residuals for most models indicate an assumption may not have been met,
some of which is the result of having a relatively small amount of data.*

First, a multiple regression for the detection time (without the construction time) with
predictors users, resources, activity count, and mean selectivity.

The most significant predictor of detection time is the mean selectivity.
```{r, echo=FALSE}
expr3.reg <- lm(detection_time_mean ~ activity_count + users + resources + selectivity_mean, expr3)
summary(expr3.reg)
ggplot(data = expr3.reg) +
  geom_point(aes(y = .resid, x = .fitted)) +
  labs(x = "Fitted Values", y = "Residuals")
ggplot(data = expr3.reg, aes(x = .resid)) +
  geom_histogram(aes(y = after_stat(density)), bins=8) +
  stat_function(fun = dnorm,
                args = list(mean = mean(expr3.reg$residuals),
                            sd = sd(expr3.reg$residuals)),
                col = "blue",
                size = 1.5) +
  labs(title="Distribution of Residuals")
```

Limitations: Since activity count was controlled, it is the most linear
variable, and thus is most likely to have a strong correlation with the detection time.
Users only increased in this experiment,
resulting in no trials with large numbers of activities and few users or resources.
Further, these variables are **not independent.**

### Adjusted (and Best) Regression for Detection Time
To address this, since the action space size is a composite of the users and resources,
these columns can be combined using the formula in the paper.
```{r, echo=FALSE}
expr3.asize <- expr3 %>% mutate(action_space_size = resources * users * (5 + 3 * users))
expr3.reg <- lm(detection_time_mean ~ activity_count + action_space_size + selectivity_mean, expr3.asize)
summary(expr3.reg)
ggplot(data = expr3.reg) +
  geom_point(aes(y = .resid, x = .fitted)) +
  labs(x = "Fitted Values", y = "Residuals")
ggplot(data = expr3.reg, aes(x = .resid)) +
  geom_histogram(aes(y = after_stat(density)), bins=8) +
  stat_function(fun = dnorm,
                args = list(mean = mean(expr3.reg$residuals),
                            sd = sd(expr3.reg$residuals)),
                col = "blue",
                size = 1.5) +
  labs(title="Distribution of Residuals")
```

### Regression for total run time
Construction time and detection time with predictors activity count, action space size, and selectivity mean.

```{r, echo=FALSE}
expr3.reg <- lm(detection_time_mean + construction_time_mean ~ activity_count + action_space_size + selectivity_mean, expr3.asize)
summary(expr3.reg)
ggplot(data = expr3.reg) +
  geom_point(aes(y = .resid, x = .fitted)) +
  labs(x = "Fitted Values", y = "Residuals")
ggplot(data = expr3.reg, aes(x = .resid)) +
  geom_histogram(aes(y = after_stat(density)), bins=8) +
  stat_function(fun = dnorm,
                args = list(mean = mean(expr3.reg$residuals),
                            sd = sd(expr3.reg$residuals)),
                col = "blue",
                size = 1.5) +
  labs(title="Distribution of Residuals")
```

### Stats
```{r, echo=FALSE}
ggplot(data = expr3) +
  geom_histogram(aes(x=detection_time_mean), bins=10) +
  theme_bw() +
  labs(title="Distribution of Detection Times")
ggplot(data = expr3) +
  geom_histogram(aes(x=detection_time_mean + construction_time_mean), bins=10) +
  theme_bw() + 
  labs(title="Distribution of Total Run Times")
```

```{r, echo=FALSE}
ggplot(data=expr3, aes(x=selectivity_level, y=selectivity_mean, fill=selectivity_level)) +
  geom_boxplot() +
  labs(title = "Variation of selectivity within a level", x = "Selectivity level", y = "Selectivity values")
```